---
title: "Coursera ML Write Up"
author: "Cory Nissen"
date: "10/23/2014"
output: html_document
---

To start the model building process, I read in the data. I noticed several of the columns had a high number of missing or blank values, so I excluded those from consideration when building the model.  

```{r eval=FALSE}
df <- read.csv("pml-training.csv", stringsAsFactors=F)
df <- df[,apply(df,2,function(x) sum(is.na(x)) / nrow(df) < .25)]
is.blank <- function(x)x==""
df <- df[,apply(df,2,function(x) sum(is.blank(x)) / nrow(df) < .25)]
```

Next, since the training set was sufficiently large, I split it up into a "training" and "test" set to get an idea of what the out of sample error might be.  
```{r eval=FALSE}
library(caret)
set.seed(1)
df$classe2 <- as.numeric(as.factor(df$classe))
df$classe3 <- as.factor(df$classe)
inTrain <- createDataPartition(y=df$classe, p=.70, list=FALSE)
df_train <- df[inTrain,]
df_test <- df[-inTrain,]
```

Then I started building models. I built the model formula as follows:  
```{r eval=FALSE}
vars <- names(df_train)[8:59]
form2 <- as.formula(paste("classe3 ~ ", paste(vars, collapse="+")))
```

I started with a tree using party::ctree. The kappa was .7993. I tried an svm model that had a kappa of .7242. I ended up using a gbm using repeated cross validation using the following code for a kappa of .9938:  
```{r eval=FALSE}
fitControl <- trainControl(method="repeatedcv",
                           number=5,
                           repeats=1,
                           verboseIter=TRUE)
gbmGrid <- expand.grid(.interaction.depth = (1:5) * 2,
                       .n.trees = (1:10)*25, .shrinkage = .1)
gbmFit <- train(form1, data=df_train,
                method="gbm", bag.fraction = 0.5,
                trControl=fitControl,
                verbose=FALSE, tuneGrid = gbmGrid)
gbmFit
gbm_pred <- predict(gbmFit, df_test)
confusionMatrix(gbm_pred, df_test$classe3)
```
The accuracy was .9954 using the "test" set. This is my estimate for out of sample error.  

Predicting using the "official" 20 test cases was 20 for 20 correct:  
```{r eval=FALSE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("output/problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
test <- read.csv("pml-testing.csv", stringsAsFactors=F)
answers <- predict(gbmFit, test)
pml_write_files(answers)
```





